---
title: Bangladesh Ground Arsenic with R
author: owenqi
date: '2019-07-05'
categories:
  - Statistics
tags:
  - R
  - ggplot2
---

This post demostrate implements the computation and visualization of CSI (Contamination Severity Index) develop by Dr. P.K Sen in 2016, data obtain from [British Geological Survey](https://www.bgs.ac.uk/arsenic/bangladesh/).

Map getting from [GADM version 2.8](https://gadm.org/download_country_v2.html)

```{r, eval=FALSE}
load('data.Rda')
thana_map <- readRDS(gzcon(url('https://biogeo.ucdavis.edu/data/gadm2.8/rds/BGD_adm3.rds')))
```

## 1. Expression

Let $Y_{(i)}, i = 1,2, \ldots, n$ be the ordered sample from a region (can be division/district/thana), in our case, **thana**. 

* Let $M$ denote the number of observation less than the threshold $L$ (given).
* Let $F_n(x) = \frac{1}{n}\sum_{i = 1}^{n}I(Y_i \leq x), x \geq 0$, the empirical distribution be the estimator of true distribution $F(x)$.

For arsenic (As) level $Y_{(i)}$, the corresponding propensity socre, denoted by $C_{ni}, i = 1,2, \ldots,n$ is given by

\begin{align}
\tag{2.1}
C_{ni} = \max(0, (1 - L/Y_i)^{\hat{\theta(Y_{(i)})}})
\end{align}

where

\begin{align}
\tag{2.2}
\hat{\theta(Y_{(i)})} = \frac{2}{n-M}\sum_{j = M+1}^{n}\frac{\min(Y_{(i)},Y_{(j)})}{Y_{(i)}+ Y_{(j)}}, i = 1,\ldots, n
\end{align}

and finally the CSI is the mean of the propensity score averaged by all $n$ observation.

\begin{align}
\tag{2.3}
\hat{C}_E = \frac{1}{n}\sum_{M \leq i \leq n}C_{ni}
\end{align}

Note: For Equation (1), if $Y_{(i)} \leq L$, then $C_{ni} = 0$.

## 2. Implementation

```{r propensity_score, eval=FALSE}
propensity_score <-function(x, y_vector, L = 10){
  # x is argument, i.e. oberservation to be convert
  # y_vector is a vector of ALL observations in the sub_region
  
  n <- length(y_vector)
  y_truncate <- y_vector[y_vector > L]

  # case that ALL observation less than the threshold
  if(length(y_truncate) == 0)
    return(rep(0, n))
  
  # when num of y_trancate == 1, then 2/(n-M-1) = infinity
  if(length(y_truncate) == 1)
    return(max(0,1-L/x))
    
  # base of Eq. (1)
  C_ni <- max(0,1-L/x)
  
  # coeff * y_sum is Eq. (2)
  coeff <- 2/length(y_truncate)
  y_sum <- sum(pmin(x,y_truncate)/(x + y_truncate))
  
  # return value of Eq. (1)
  return(C_ni^(coeff*y_sum))
}
```

Once we have the CSI score for the thana, CSI is simply the mean of the scores of **all observations**.
```{r csi, eval=FALSE}
csi <- function(y_vector){
  # find propensity score for each observation
  score <- sapply(y_vector, y_vector = y_vector, propensity_score)
  # mean of the score is csi
  return(mean(score))
}
```

## 3. Result

We present two type of plot, 

* The first one we use the exact **CSI values** value computed from the estimator, it is presented in a **continuous scale**.

```{r data_merge, message=FALSE, warning=FALSE, eval=FALSE}
library(tibble)
library(ggplot2)
library(dplyr)
thana_obs <- split(dataset, dataset$thana_id)

thana_csi <- function(data_list, discrete = FALSE){
  n <- length(data_list)
  csi_vector <- rep(NA, n)
  ids <- names(data_list)
  for(i in 1: n){
  csi_vector[i] <- csi(thana_obs[[i]]$as)
  }
  cutoff <- seq(0, 1, length.out = 11)
  if (discrete == TRUE){
    csi_vector <- cut(csi_vector, breaks = cutoff, include.lowest = TRUE)
    csi_vector <- factor(csi_vector, levels = rev(levels(csi_vector)))
    }
  # can consider rounding to 3 decimals
  df <- tibble(id = ids, csi = csi_vector)
  return(df)
}

# csi in dataframe
csi_df <- thana_csi(thana_obs)
# plotting result
thana_df <- fortify(thana_map)
# merge result
plot_csi <- left_join(thana_df, csi_df, by = 'id')

dcsi_df <- thana_csi(thana_obs, discrete = TRUE)
plot_dcsi <- left_join(thana_df, dcsi_df, by = 'id')
```

```{r, out.width="150%", echo=FALSE, message=FALSE}
load('data/csi.Rda')
thana_map <- readRDS('maps/BGD_adm3.Rds')
library(ggplot2)

p_csi <- ggplot(plot_csi, aes(x = long, y = lat, group = group, fill = csi)) + geom_polygon(color = "white") +
  scale_fill_viridis_c() + 
  coord_map() +  ggtitle("CSI based on Empirical Distribution") +
  xlab("Longitude") + ylab("Latitude") +  labs(fill = "CSI") + theme_minimal() +
  guides(fill = guide_colourbar(barwidth = 1, barheight = 18)) + 
  theme(plot.title = element_text(hjust = 0.5))
p_csi
```

* The second one we use the exact 10 bins, from which we divide 0 - 1 into 10 categories, it is presented in a **discrete scale**.

```{r, out.width="150%", echo=FALSE, message=FALSE}
p_dcsi <- ggplot(plot_csi, aes(x = long, y = lat, group = group, fill = dcsi)) + geom_polygon(color = "white") +
  scale_fill_viridis_d(direction = -1, na.value = "grey50") + 
  coord_map() +  xlab("Longitude") + ylab("Latitude") +  labs(fill = "CSI") + theme_minimal() +
  ggtitle("CSI based on Empirical Distribution") +
  theme(plot.title = element_text(hjust = 0.5))
p_dcsi
```

## References

1. Sen, P. K. (2016). Abundant Environmental Arsenic Contamination: Some Statistical Perspectives. Sankhya B, 78(2), 341-361.

3. [Viridis Color Scales](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html)